Future Cosmic Microwave Background delensing with galaxies surveys.

### 2017-05-29 Notes/DES_rec_phi_paper.tex
order

The measurement of the cosmic microwave background (CMB) polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe. 
A particular component, the so-called B-modes is indeed a direct signature of the presence of gravitational waves in the early universe. However improving the level of noise is not enough. 
This is even truer if the aim is to not only detect the amplitude of gravitational waves but also the shape of their spectrum to test for example inflation consistency relations.
Removing the lensing component from the measurement of CMB B-modes will be important to constrain the amplitude of the primordial gravitational wave contribution to the signal.
Here we discuss the role of current and future large scale structure surveys in improving the reconstruction of the lensing potential that lenses the CMB photons and how this reflects into an improved delensing efficiency.
We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$)
We find that..
\end{abstract}

@@

The amplitude of this primordial B-mode component is parametrized by the ratio of the amplitudes of the primordial tensor and scalar perturbations, $r$. The shape of this component, i.e. the amplitude as a function of the angular scales is encoded in the tensor tilt $n_{t}$.
Most of the inflationary models directly relate the value of r to the energy scale of inflation.
CMB B modes provide the cleanest known observational window into the primordial gravitational waves background.
Measuring or constraint the value of $r$ and $n_t$ will put tight constraint on the allowed models of inflation.

@@

However reducing the level of noise will not be enough to reach this goal.
Unfortunately the observed B modes are not solely due primordial gravitational waves but also by late time effects. In this work we will focus on gravitational interactions with LSS.
Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,polarbear2014c,vanengelen14b,planck2015XV} and directly from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.

This component acts as a source of confusion  for searches of the PGW background.
Indeed, current experiment are already limited by the contamination from lensing B modes since the instrument noise is below the lensing B mode $rms$ fluctuations \citep{bk14}.
The optimal way to approach this source of noise is to build the specific realization of lensing B modes on the sky can be and used it to clean the data in a process called ``delensing''.

Delensing has been studied for many years \cite{knox2002, kesden2002, seljak2003, simard:2015,sherwin15,smith:2012}.
Furthermore recently it has been performed on CMB temperature data using the cosmic infrared background in \cite{larsen:2016}, on CMB temperature and polarization data using internal CMB reconstruction \cite{ carron17}. Finally the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in  where $~28\%$ of the lensing power has been removed.
\vspace{0.5cm}

In this paper we

\vspace{0.5cm}

This paper is organized as follow: in \refsec{th} we describe the lensing B-mode components and the residual power after delensing with tracers of the lensing potential. We describe the LSS tracers used in this analysis in \refsec{model}. \refsec{for} is the main results of this work: the improvement to inflationary parameters due to delensing with CMB and LSS. We conclude in \refsec{concl}.

@@

As for the CMB photons temperature, the intensity map of photons on the sky, also the Q and U mode decomposition of their polarization is remapped by lensing as:
\be
Q(\hat{\mathbf{n}}) = Q_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi);~~
U(\hat{\mathbf{n}}) = U_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi)
\ee
where the deflection angle is the gradient of lensing potential potential integrated along the line of sight $\nabla\phi$.
The CMB polarization is usually decomposed into odd-parity Fourier modes E and B. 
As shown in \cite{} this for symmetry reason tensor perturbations are the main source of the B-modes configuration. For this reason, B-modes are a promising signature of early universe tensor perturbations.
However, this promising sign of primordial gravitational waves is partially obscured by a secondary mechanism that can produce B-modes also in the late universe. Indeed the gravitational interaction with large scale structure generates CMB B-modes by distorting primordial E-modes.

@@

The full B-mode power spectrum measured on the sky is given by a possible primordial component $C_l^{BB,r}$ together with the lensing $C_l^{BB,\mathrm{lens}}$ contribution and the instrumental noise $N_l^{BB}$:
\be
C_l^{BB,\mathrm{full}} = C_l^{BB,r} + C_l^{BB,\mathrm{lens}} + N_l^{BB}.
\ee
Given the current constraints, the lensing component is a significant source of B-modes that, for GW searches purposes correspond to a white noise source of roughly $5 \mu K$-arcmin. This means that it is not only bigger than the biggest allowed GW contribution at scales bigger than several degrees but it is comparable with the decreasing level of instrumental noise. 
For this reason, it is critical to characterize and eventually remove it from the data.
To do so, we need to build a template \refeq{blens} (or its real-space analog) of the actual lensing modes given a measurement of the E-mode field and the lensing potential $\phi$.
While E is measured directly, we ca estimate $\phi$ using "tracers" of the dark matter distribution that creates the potential.

@@

The bigger $\rho_l$ is for a LSS field the more it is correlated with the lensing potential acting on the CMB photons. An higher correlation allows for a better reconstruction of the $\phi^{CMB}$ and, as a consequence, of $B^{lens}$.

@@

Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing but it also depends on how much it is correlated with the already used set of tracers.
\reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
Galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST,DES and DESI curves.
However given the significant low noise of this measurement they can still help the delensing of the CMB.
Furthermore, the low-redshift tail of the lensing kernel is also the one affected by large scale structures non-linearities. They can indeed be extremely helpful if future CMB surveys will try to remove that contribution instead of attempting to model it.
On the other end, CIB and 21 cm surveys are probing to high redshift structures and independently from the model assumed they show a fairly good overlap with the CMB lensing kernel.

@@

The CMB lensing potential is \textit{the} field that we need to reconstruct to reverse the effect of large scale structure and delens the CMB.
However, the lensing potential can also be reconstructed from the CMB itself. In that case, we can treat it as a noisy tracer of the true field. Both the CMB lensing fields and its noisy reconstructed counterparts have the same kernel $W^{\kappa}(z)$.
However, when computing the power spectrum of the latter, we need to add a noise component. Given the instrumental noise level and the beam, we can compute the reconstruction noise $N_l^{\kappa \kappa}$, and so
\beq
\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_l
\eeq
In this work, the level of noise is computed assuming an iterative quadratic estimator is used in the reconstruction

@@

\subsection{Cosmic infrared Background (CIB)}
\label{sec:cib}
Following \cite{Addison:2011se}, we model the CIB power directly as $C_{\ell}^{{\rm CIB}\mbox{-}{\rm CIB}} = 3500 (l/3000)^{-1.25} {\rm Jy^2 / sr}$.
This model provides an accurate fit to several experimental results.
For the cross-spectra with the CMB lensing or other galaxies tracers $C_{\ell}^{{\rm CIB}\mbox{-}j}$, we use the single-SED model of~\cite{hall10}.
It corresponds to the kernel:
\begin{align}\label{eq:12}
W^{\text{CIB}}(z) = b_c\ \frac{\chi^2(z)}{H(z)(1+z)^2}\ e^{-\frac{(z-z_c)^2}{2\sigma^2_z} } f_{\nu(1+z)},
\end {align}
for
\begin{equation}
f_{\nu} = 
\begin{cases}
\Big( e^{\frac{h\nu}{kT}} - 1 \Big)^{-1} \nu^{\beta+3} & (\nu \leq v^{\prime}) \\ \Big( e^{\frac{h\nu^{\prime}}{kT}} - 1 \Big)^{-1} \nu^{\prime \beta+3} \Big( \frac{\nu}{\nu^{\prime}} \Big)^{-\alpha} & (\nu > v^{\prime})
\end{cases}
\end{equation}

@@

\section{Parameter constraints improvement after delensing}
\label{sec:for}
In this section, we forecast the expected delensing efficiency and the relative importance of galaxies tracers for delensing current and future experiments. We will use a Fisher method and measured the delensign efficiency by the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
We assume a CMB experimental scenario composed of a high-resolution CMB experiments which define the internal reconstruction efficiency, together with a low-noise, low-resolution experiment which B-modes will be delensed and used to constrain the inflationary B-modes.
We focus on three different scenarios: the current stage, the 3G one and the futuristic CMB Stage 4.

\subsection{Fisher method}
In this framework, the error on the cosmological parameters constrained by a B-mode spectrum measurement is given as usual by:
\begin{equation}
\label{eqn:fisher}
\sigma(p)= \left[\sum_l^{\ell_{\rm{max}}} \left(\frac{\partial {C}_{\ell}^{BB,r,n_{t}}}{\partial p}\right)^2/\mathrm{Cov}^2( C_{\ell}^{BB} ) \right]^{-\frac{1}{2}}
\end{equation}
where we consider the parameters $p =\{r,n_{t}\}$. 
We will assume a gaussian covariance:
\beq
\sigma( C_{l}^{BB} )= \sqrt{ \frac{2}{(2 {l}+1) \fsky }} \left( C_l^{BB,r} + C_{{l}}^{BB,\mathrm{lens}}+N_{{l}}^{BB} \right).
\eeq
Here the B-modes noise spectrum is given by:
\beq
N_l^{BB} = \left({\Delta_P}/{T_{\mathrm{CMB}}}\right)^2  e^{ {l^2 \theta_\mathrm{FWHM}^2}/({8 \ln 2})}
\eeq
where $\theta_\mathrm{FWHM}$ is the full half width of the telescope  beam and $\Delta_P$ is the instrumental noise of the experiment.
Even if the covariance structure has known non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012} this approximation is good enough to show the improvement due to delensing.

It can be seen from \refeq{fisher}, that the parameter constraints are inversely proportional to the covariance of the measurement. Since lensing B-mode $C_{{l}}^{BB,\mathrm{lens}}$ are a substantial component of the covariance, removing part of them trough delensing will improve parameter constraints.

Note that we are fixing the other cosmological parameters. Uncertainties in those will however propagate to uncertainties in $\{r,n_{t}\}$. While the constraints will be slightly optimistic, this has no real impact on the estimate of the improvement due to delensing which is the main focus of this work.
We define the improvement due to delensing as the ration of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$
%===================
\subsection{Delensing with current CMB and LSS}

@@

Very recently delensing has been proven possible on real data using as tracers both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as the CIB \cite{larsen:2016a,manzotti:2017}.
Here we discuss the improvement due to combining these and other available probes (see \cite{yu:2017} for publicly averrable multi-tracers data-products).

On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with a noise equal to \textcolor{red}{$3\mu$K-arcmin} and a beam of \textcolor{red}{30 arcmin}.
The CMB lensing reconstruction will be performed by a higher resolution CMB experiment. We explore two possibilities.
First, we test the efficiency with internal reconstruction performed by Planck \cite{2016A&A...594A..15P} using the CMB lensing noise curves publicly available
\footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
The second one uses a ground-based experiment with the final noise level of SPT-Pol which correspond to \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.

@@

The correlation attainable using current experiments is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol and.
In general, we observe that CMB lensing reconstruction correlate very well at low $\ell$  and it then falls rapidly because of the the raise of the reconstruction noise.
For the noise levels of Planck, the internal reconstruction is at most $70\%$ correlated at very large scales end falls rapidly to $40\%$ at $\ell=200$. Its correlation is at the level of the one coming from LSS at almost all scales.
On the other end, the CMB lensing reconstruction \cite{} from SPT-Pol will be more than 80\% correlated with the true field at large scales $\ell<300$ where the correlation become comparable with the current LSS surveys.

Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by very large Galactic dust residuals.
Optical surveys galaxies like DES are less affected by that systematics and can be used at larger scales.
Furthermore, in \reffig{corrnow} DES contribution to delensing is higher than WISE because the redshift distributions peaks at slightly higher redshift thus overlapping better with lensing.

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{../images/actual_scenario_planck.pdf}
\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the multipole $\ell$. }
\label{fig:corrnow-planck}
\end{center}
\end{figure}

@@

Using this correlation level we can, using \refeq{Bres} compute the residual B-mode power after delensing and test the consequent improvement on parameter constraints with \refeq{fisher}. 
For current surveys we only consider $r$ since the broad beam of the deep BK-like survey does not allow enough leverage to constrain the shape of the B-mode spectra.
We first compute the improvement with a fiducial value of $r_{\rm{fid}}=0$.
The results are summarized in \reftab{current}

@@

The results are in agreement with \cite{yu:2017,larsen:2016a}, with the CIB being the predominant source of delensing with an improvement of almost 50\%. Planck and DES are essential to reconstruct the largest scales of the lensing potential and as such the play an important role in delensing. The improvement due to WISE is quite marginal, but given the big sky coverage, it might be useful to delens large CMB polarization experiments.
For reference we also test a scenario where GW are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}. As expected the importance of delensing itself is now reduced given that the lensing component constitutes a smaller portion of the total B-modes variance.

@@

 \begin{table}
\caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to ration of the error before and after delensing.}
  \vspace{0.2cm}
  \begin{tabular}{ | c | c | c|}
\hline
Surveys & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
WISE & 1.10 &  \\ \hline
DES & 1.21  & \\ \hline
CIB & 1.52  & \\ \hline
LSS & 1.78  & \\ \hline
CMB & 1.62  &  \\ \hline
LSS+CMB & 2.35 &   \\ \hline

@@

\subsection{CMB-S3 Era}
The accuracy of CMB is rapidly improving and the next generation ground based telescopes have been already deployed and are currently taking data.
For the CMB at the beginning of the next decade, we will assume a deep experiment with level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam. This is a reasonable level for a combination of the next generation of BK and SPT3G. This level of noise will also be assumed for the internal noise reconstruction.

The correlation attainable using generation 3 experiments is shown in \reffig{corrS3}.
Not only will CMB experiments improved in the near future: possibly D.E.S.I will start taking data. For this reason we add DESI to the previously mentioned LSS tracers. 
DESI will be slightly more efficient than a DES-like surveys because of the slightly higher redshift that it is able to probe.
However the CIB will still play the major role among LSS.

@@

An ambitious program for a generation 4 ground CMB experiment is currently under planning. Moreover satellite experiments have been proposed.
The correlation attainable using current experiments is shown in \reffig{corrs4},
In this very futuristic but reasonable scenario we add to the avaialable large scale structures both 21 cm surveys like SKA and large scale structures like LSST and Euclid.

@@

\subsection{Bias uncertainties degradation}
The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of inflationary constraint of delensed spectra.
In this section we quantify this effect.
We will now marginalize over unknown galaxies parameters but we will use a full dataset of CMB and galaxies data. The idea is that as shown in the low level of noise in galaxies surveys might allow us to internally calibrate them.
We will use a Fisher approach the Fisher matrix is:
\beqn
F_{pq} &=& \sum_{l_{a} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}  \sum_{l_{b} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}   \frac{\partial {C}_{l_a}^{BB,\mathrm{del}}}{\partial \theta_p } \left[ \mathrm{Cov}^{BB,BB}\right]_{l_a, l_b} ^{-1} \frac{\partial {C}_{l_b}^{BB,\mathrm{del}}}{\partial \theta_q }\nonumber \\&+&  \sum_j \frac{\frac{\partial {C}_j^{\kappa I}}{\partial \theta_p } \frac{\partial {C}_j^{\kappa I}}{\partial \theta_q }}{(\Delta C_j^{\kappa I})^2}
+\sum_j \frac{\frac{\partial {C}_j^{I I}}{\partial \theta_p } \frac{\partial {C}_j^{I I}}{\partial \theta_q }}{(\Delta C_j^{I  I})^2}
\eeqn

>>>>

The measurement of the cosmic microwave background (CMB) polarization is a promising experimental dataset to test the inflationary paradigm and to probe the physics of the early universe. 
A particular component, the so-called B-modes is indeed a direct signature of the presence of gravitational waves in the early universe. However improving the level of noise is not enough. 
This is even truer if the aim is to not only detect the amplitude of gravitational waves but also the shape of their spectrum to test for example inflation consistency relations.
Removing the lensing component from the measurement of CMB B-modes will be important to constrain the amplitude of the primordial gravitational wave contribution to the signal.
Here we discuss the role of current and future large scale structure surveys in improving the reconstruction of the lensing potential that lenses the CMB photons and how this reflects into an improved delensing efficiency.
We quantify this by the improvement due to delensing on the constraints on the inflationary tensor perturbations amplitude and shape ($r$ and $n_{t}$)
We find that current generation experiments...
\end{abstract}

@@

The amplitude of this primordial B-mode component is parameterized by the ratio of the amplitudes of the primordial tensor and scalar perturbations, $r$. The shape of this component, i.e. the amplitude as a function of the angular scales is encoded in the tensor tilt $n_{t}$.
Most of the inflationary models directly relate the value of r to the energy scale of inflation.
CMB B-modes provide the cleanest known observational window into the primordial gravitational waves background.
Measuring or constraint the value of $r$ and $n_t$ will put a tight constraint on the allowed models of inflation.

@@

However reducing the level of noise will not be enough to reach this goal.
Unfortunately, the observed B modes are not solely due to primordial gravitational waves but also by late time effects. In this work, we will focus on gravitational interactions with LSS.
Lensing shears the CMB polarization pattern, producing ``lensing B modes'' from CMB E modes \citep{zaldarriaga98}.
This expected component has now been measured both in cross-correlation with LSS \citep{hanson13,polarbear2014c,vanengelen14b,planck2015XV} and directly from CMB data alone \citep{polarbear2014b,bicep2a,keisler15}.

This component acts as a source of confusion for searches of the PGW background.
Indeed, current experiments are already limited by the contamination from lensing B modes since the instrument noise is below the lensing B-mode $rms$ fluctuations \citep{bk14}.
The optimal way to approach this source of noise is to build the specific realization of lensing B modes on the sky can be and used it to clean the data in a process called ``delensing''.

Delensing has been studied for many years \cite{knox2002, kesden2002, seljak2003, simard:2015,sherwin15,smith:2012}.
Furthermore recently it has been performed on CMB temperature data using the cosmic infrared background in \cite{larsen:2016}, on CMB temperature and polarization data using internal CMB reconstruction \cite{ carron17}. Finally, the highest B-mode delensing efficiency has been achieved with SPT and Herschel data in  where $~28\%$ of the lensing power has been removed.

In this paper, we study the possible role of future galaxies surveys as tracers of the lensing potential. These will translate into a better reconstruction of the B-modes in the measured patch and as a consequence improve the constraints on inflation trough delensing. We model several actual and future surveys, and after computing the residual B-modes, we forecast the statistical uncertainness on $r$ and $n_T$ for current CMB experiments as well as the next generation (S3) and the planned fourth generation (S4).

This article is organized as follow: in \refsec{th} we describe the lensing B-mode components and the residual power after delensing with tracers of the lensing potential. We describe the LSS tracers used in this analysis in \refsec{model}. \refsec{for} is the main results of this work: the improvement of inflationary parameters constraints due to delensing with CMB and LSS. We conclude in \refsec{concl}.

@@

As for the CMB photons temperature, the intensity map of photons on the sky, also the Q and U mode decomposition of their polarization is remapped by lensing as:
\be
Q(\hat{\mathbf{n}}) = Q_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi);~~
U(\hat{\mathbf{n}}) = U_{\mathrm{unlensed}}(\hat{\mathbf{n}}+\nabla\phi)
\ee
where the deflection angle is the gradient of lensing potential potential integrated along the line of sight $\nabla\phi$.
The CMB polarization is usually decomposed into odd-parity Fourier modes E and B. 
As shown in \cite{} this for symmetry reason tensor perturbations are the principal source of the B-modes configuration. For this reason, B-modes are a promising signature of early universe tensor perturbations.
However, this promising sign of primordial gravitational waves is partially obscured by a secondary mechanism that can produce B-modes also in the late universe. Indeed the gravitational interaction with large scale structure generates CMB B-modes by distorting primordial E-modes.

@@

The full B-mode power spectrum measured on the sky is given by a possible primordial component $C_l^{BB,r}$ together with the lensing $C_l^{BB,\mathrm{lens}}$ contribution and the instrumental noise $N_l^{BB}$:
\be
C_l^{BB,\mathrm{full}} = C_l^{BB,r} + C_l^{BB,\mathrm{lens}} + N_l^{BB}.
\ee
Given the current constraints, the lensing component is a significant source of B-modes that, for GW searches purposes correspond to a white noise source of roughly $5 \mu K$-arcmin. This means that it is not only bigger than the biggest allowed GW contribution at scales bigger than several degrees, but it is comparable with the decreasing level of instrumental noise. 
For this reason, it is critical to characterize and eventually remove it from the data.
To do so, we need to build a template \refeq{blens} (or its real-space analog) of the actual lensing modes given a measurement of the E-mode field and the lensing potential $\phi$.
While E is measured directly, we ca estimate $\phi$ using "tracers" of the dark matter distribution that creates the potential.

@@

The bigger $\rho_l$ is for an LSS field, the more it is correlated with the lensing potential acting on the CMB photons. A higher correlation allows for a better reconstruction of the $\phi^{CMB}$ and, as a consequence, of $B^{lens}$.

@@

Note that the gain in adding a new tracer is not only proportional to its correlation with the CMB lensing, but it also depends on how much it is correlated with the already used set of tracers.
\reffig{kernel} show the different kernels as a function of redshift computed using the models and parameters described in \refsec{model}.
Galaxies clustering surveys can only reconstruct the low-z portion of the lensing kernel as can be seen from the LSST, DES and DESI curves.
However given the significant low noise of this measurement they can still help the delensing of the CMB.
Furthermore, the low-redshift tail of the lensing kernel is also the one affected by large-scale structures non-linearities. They can indeed be extremely helpful if future CMB surveys will try to remove that contribution instead of attempting to model it.
On the other end, CIB and 21 cm surveys are probing to high redshift structures and independently from the model assumed they show a relatively good overlap with the CMB lensing kernel.

@@

The CMB lensing potential is \textit{the} field that we need to reconstruct to reverse the effect of large scale structure and delens the CMB.
However, the lensing potential can also be reconstructed using the CMB itself. In that case, we can treat it as a noisy tracer of the true field. Both the CMB lensing fields and its noisy reconstructed counterparts have the same kernel $W^{\kappa}(z)$.
However, when computing the power spectrum of the latter, we need to add a noise component. Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_l^{\kappa \kappa}$, and so
\beq
\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_l
\eeq
In this work, the level of noise is computed assuming an iterative quadratic estimator is used in the reconstruction

@@

\subsection{Cosmic infrared Background (CIB)}
\label{sec:cib}
Following \cite{Addison:2011se}, we model the CIB power directly as $C_{\ell}^{{\rm CIB}\mbox{-}{\rm CIB}} = 3500 (l/3000)^{-1.25} {\rm Jy^2 / sr}$.
This model provides an accurate fit to several experimental results.
For the cross-spectra with the CMB lensing or other galaxies tracers, $C_{\ell}^{{\rm CIB}\mbox{-}j}$, we use the single-SED model of~\cite{hall10}.
It corresponds to the kernel:
\begin{align}\label{eq:12}
W^{\text{CIB}}(z) = b_c\ \frac{\chi^2(z)}{H(z)(1+z)^2}\ e^{-\frac{(z-z_c)^2}{2\sigma^2_z} } f_{\nu(1+z)},
\end {align}
for
\begin{equation}
f_{\nu} = 
\begin{cases}
\Big( e^{\frac{h\nu}{kT}} - 1 \Big)^{-1} \nu^{\beta+3} & (\nu \leq v^{\prime}) \\ \Big( e^{\frac{h\nu^{\prime}}{kT}} - 1 \Big)^{-1} \nu^{\prime \beta+3} \Big( \frac{\nu}{\nu^{\prime}} \Big)^{-\alpha} & (\nu > v^{\prime})
\end{cases}
\end{equation}

@@

\section{Parameter constraints improvement after delensing}
\label{sec:for}
In this section, we forecast the expected delensing efficiency and the relative importance of galaxies tracers for delensing in current and future experiments. We will use the Fisher information matrix to quantify the delensign efficiency as the improvement in the constraint of two inflationary parameters: the tensor to scalar ratio $r$ and the tensor tilt $n_{T}$.
We assume a CMB experimental scenario composed of a high-resolution CMB experiment which defines the internal reconstruction efficiency, together with a low-noise, low-resolution experiment which B-modes will be delensed and used to constrain the inflationary parameters.
We focus on three different scenarios: the current stage, a third generation stage (3G)  and finally the futuristic CMB Stage 4.

\subsection{Fisher method}
In this framework, the error on a cosmological parameter constrained by a B-mode spectrum measurement is given as usual by:
\begin{equation}
\label{eqn:fisher}
\sigma(p)= \left[\sum_l^{\ell_{\rm{max}}} \left(\frac{\partial {C}_{\ell}^{BB,r,n_{t}}}{\partial p}\right)^2/\mathrm{Cov}^2( C_{\ell}^{BB} ) \right]^{-\frac{1}{2}}
\end{equation}
where we consider the parameters $p =\{r,n_{t}\}$. 
We will assume a gaussian covariance:
\beq
\sigma( C_{l}^{BB} )= \sqrt{ \frac{2}{(2 {l}+1) \fsky }} \left( C_l^{BB,r} + C_{{l}}^{BB,\mathrm{lens}}+N_{{l}}^{BB} \right).
\eeq
Here the B-modes noise spectrum is given by:
\beq
N_l^{BB} = \left({\Delta_P}/{T_{\mathrm{CMB}}}\right)^2  e^{ {l^2 \theta_\mathrm{FWHM}^2}/({8 \ln 2})}
\eeq
where $\theta_\mathrm{FWHM}$ is the full half width of the telescope beam and $\Delta_P$ is the instrumental noise of the experiment.
Even if the covariance structure has non-Gaussian contributions \cite{motloch:2017,benoit-levy:2012} this approximation is good enough to show the improvement due to delensing.

The reason why removing the lensing contribution will improve parameters constraints can be seen from \refeq{fisher}.The parameter uncertainties are inversely proportional to the covariance of the measurement. Since lensing B-mode $C_{{l}}^{BB,\mathrm{lens}}$ is a substantial component of the covariance, removing part of them trough delensing will improve parameter constraints.

Note that we are fixing the other cosmological parameters in \refeq{fisher}. Uncertainties in those will however propagate to uncertainties in $\{r,n_{t}\}$. While the constraints will be slightly optimistic, this has no real impact on the estimate of the improvement due to delensing which is the main focus of this work.
We define that improvement as the ration of the constraints before and after delensing: $\alpha_{r} = \sigma^{\rm{del}}(r)/\sigma(r)$ and $\alpha_{n_{t}} = \sigma^{\rm{del}}(n_{t})/\sigma(n_{t})$
%===================
\subsection{Delensing with current CMB and LSS}

@@

Very recently delensing has been proven possible on real data using as tracers both the internal CMB lensing potential reconstruction \cite{carron:2017} as well as the CIB \cite{larsen:2016a,manzotti:2017}.
Here we discuss the improvement due to combining these and other available probes (see \cite{yu:2017} for publicly accessible multi-tracers data-products).

On the CMB side, we will combine a Bicep-Keck like \textit{deep} CMB experiment with a noise equal to \textcolor{red}{$3\mu$K-arcmin} and a beam of \textcolor{red}{30 arcmin}.
The CMB lensing reconstruction will be performed by a higher resolution CMB experiment. We explore two possibilities.
First, we test the efficiency with an internal reconstruction performed by Planck \cite{2016A&A...594A..15P} using the CMB lensing noise curves publicly available
\footnote{\url{https://wiki.cosmos.esa.int/planckpla2015/index.php/Specially\_processed\_maps}}.
Then we will assume a ground-based experiment with the final noise level of SPT-Pol which corresponds to \textcolor{red}{$9.4\mu$K-arcmin} in polarization with a beam equal to \textcolor{red}{1.2 arcmin}.

@@

The correlation attainable using current experiments is shown in \reffig{corrnow-planck} for Planck and in \reffig{corrnow} for SPT-Pol.
In general, we observe that CMB lensing reconstruction correlate very well at low $\ell$  and it then falls rapidly because of the raise of the reconstruction noise.
For the noise levels of Planck, the internal reconstruction is at most $70\%$ correlated at very large scales and falls rapidly to $40\%$ at $\ell=200$. Its correlation is comparable to the on of LSS at almost all scales.
On the other end, the CMB lensing reconstruction \cite{} from SPT-Pol will be more than 80\% correlated with the true field at large scales $\ell<300$, where the correlation become comparable with the current LSS surveys.

Following \cite{yu:2017} we cut both the CIB and WISE at $\ell<100$ where they are contaminated by large Galactic dust residuals.
Optical surveys galaxies like DES are less affected by that systematics and can be used at larger scales.
Furthermore, in \reffig{corrnow} DES contribution to delensing is higher than WISE because the redshift distributions peaks at slightly higher redshift thus overlapping better with lensing.

\begin{figure}[htbp]
\begin{center}
\includegraphics[scale=1.]{../images/actual_scenario_planck.pdf}
\caption{Correlation factor between current galaxies survey and internally reconstructed  $\phi$ CMB lensing potential as a function of the angular scale $\ell$. 
The dashed curve corresponds to (with arbitrary scale) $C^{\kappa \kappa} \times \frac{\partial C^{BB}}{\partial{C^{\kappa \kappa}}}$, showing which scales contribute to the $\ell<100$ B-mode power.}
\label{fig:corrnow-planck}
\end{center}
\end{figure}

@@

Using this correlation level we can, using \refeq{Bres} compute the residual B-mode power after delensing and test the consequent improvement on parameter constraints with \refeq{fisher}. 
For current surveys we only consider $r$ since the broad beam of the deep BK-like survey does not provide enough leverage to constrain the shape of the B-mode spectra.
We first compute the improvement with a fiducial value of $r_{\rm{fid}}=0$.
The results are summarized in \reftab{current}

@@

When the same tracers are used, the results are in agreement with \cite{yu:2017,larsen:2016a}, with the CIB being the predominant source of delensing with an improvement of almost 50\%. Planck and DES are essential to reconstruct the largest scales of the lensing potential and as such the play a major role in delensing. The improvement due to WISE is quite marginal, but given the big sky coverage, it might be useful to delens large CMB polarization experiments.
For reference we also test a scenario where GW are present at their highest possible value of $r_{\rm{fid}}=0.12$ \cite{bicep2/keck-collaboration:2015}. As expected the importance of delensing itself is now reduced given that the lensing component constitutes a smaller portion of the total B-modes variance.

@@

 \begin{table}
\caption{$\alpha(r)$: Current generation improvements on r due to delensing. Values correspond to the ratio of the error before and after delensing.}
  \vspace{0.2cm}
  \begin{tabular}{ | c | c | c|}
\hline
Surveys & $\alpha(r=0)$ & $\alpha(r=0.12)$ \\ \hline \hline
WISE & 1.10 &  \\ \hline
DES & 1.21  & \\ \hline
CIB & 1.52  & \\ \hline
LSS & 1.78  & \\ \hline
CMB & 1.62  &  \\ \hline
LSS+CMB & 2.35 &   \\ \hline

@@

\subsection{CMB-S3 Era}
CMB measurements are rapidly improving, and the next generation of ground-based telescopes has been already deployed, and data are currently being taken.
For the CMB at the beginning of the next decade, we will assume a deep experiment with a level of noise of \textcolor{red}{$3\mu$K-arcmin} with a \textcolor{red}{1 arcmin} beam. This is a reasonable level for a combination of the next generation of BK and SPT3G. This level of noise will also be assumed for the internal noise reconstruction.

The correlation attainable using generation 3 experiments is shown in \reffig{corrS3}.
Not only will CMB experiments improved shortly: possibly D.E.S.I will start taking data. For this reason, we add DESI to the previously mentioned LSS tracers. 
DESI will be slightly more efficient than DES-like survey because of the slightly higher redshift that it can probe.
However, the CIB will still play the major role among LSS.

@@

An ambitious program for a generation 4 ground CMB experiment is currently under planning. Moreover, satellite experiments have been proposed.
The correlation attainable using current experiments is shown in \reffig{corrs4},
In this very futuristic but reasonable scenario, we add to the available large scale structures both 21 cm surveys like SKA and large scale structures like LSST and Euclid.

@@

\subsection{Bias uncertainties degradation}
The uncertainties in the theoretical assumptions used to model the galaxies can cause a degradation of the improvement of the inflationary constraint of delensed spectra.
In this section, we quantify this effect.
We will now marginalize over unknown galaxies parameters, but we will use a full dataset of CMB and galaxies data. The idea is that as shown in the low level of noise in galaxies surveys might allow us to calibrate them internally.
We will use a Fisher approach the Fisher matrix is:
\beqn
F_{pq} &=& \sum_{l_{a} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}  \sum_{l_{b} =l^{BB}_{\mathrm{min}}}^{l^{BB}_{\mathrm{max}}}   \frac{\partial {C}_{l_a}^{BB,\mathrm{del}}}{\partial \theta_p } \left[ \mathrm{Cov}^{BB,BB}\right]_{l_a, l_b} ^{-1} \frac{\partial {C}_{l_b}^{BB,\mathrm{del}}}{\partial \theta_q }\nonumber \\&+&  \sum_j \frac{\frac{\partial {C}_j^{\kappa I}}{\partial \theta_p } \frac{\partial {C}_j^{\kappa I}}{\partial \theta_q }}{(\Delta C_j^{\kappa I})^2}
+\sum_j \frac{\frac{\partial {C}_j^{I I}}{\partial \theta_p } \frac{\partial {C}_j^{I I}}{\partial \theta_q }}{(\Delta C_j^{I  I})^2}
\eeqn
